Relative parabolic factorizations of Chevalley groups

### 2015-04-12 stability-conditions.tex
Absolute stable rank rewritten + typesetting fixes

\subsection{Stable rank}
Recall the definition of {\it stable rank} introduced by H.~Bass and L.~Vaserstein (see~\cite{Ba64, Va69}).
\begin{dfn}
Recall that a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ is called {\it $I$-unimodular} if elements $a_1-1,a_2,\ldots,a_n$ are contained in $I$ while $a_1,a_2,\ldots,a_n$ generate $R$ as an ideal.
\end{dfn}
A column $b \in R^n$ will be called $I$-unimodular if its transpose $b^t$ is an $I$-unimodular row. We denote the set of all $I$-unimodular rows (resp. columns) by $\Umd(n,R,I)$ (resp. $\Ums(n,R,I)$).
When $R=I$ we refer to $R$-unimodular rows and columns as unimodular.

\begin{lemma}\label{lemma:relstrlemma}
For any $I$-unimodular row $a=(a_1,\ldots,a_n)$ there exists $I$-unimodular column $b=(b_1,\ldots,b_n)^t$ such that $ab=\sum\limits_{i=1}^n a_i b_i=1$.
\end{lemma}

\begin {dfn} An $I$-unimodular row $a=(a_1,\ldots a_{n+1})$ is called {\it stable} if one can choose $b_1,\ldots,b_n\in R$ such that the row $(a_1+a_{n+1}b_1,\ldots,a_n+a_{n+1}b_n)$ is also $I$-unimodular. If, moreover, elements $b_1,\ldots,b_n$ can be chosen from  $I$ then such $a$ will be called {\it $I$-stable}. \end{dfn}

We say that a pair $(R,I)$ satisfies stable range condition $\SR_n(R,I)$ if any $I$-unimodular row of length $n+1$ is stable.
\begin{lemma}\label{lemma:relstrlemma2}\strut
\begin{enumerate}
\item The condition $\SR_n(R,I)$ implies $\SR_m(R,I)$ for any $m\geqslant n$.
\item The condition $\SR_n(R,I)$ implies that any $I$-unimodular row of length $n+1$ is $I$-stable.
\end{enumerate}
\end{lemma}

By definition, the {\it relative stable rank} $\sr(R,I)$ of a pair $(R,I)$ is the smallest natural number $n$ such that condition $\SR_n(R,I)$ holds.
If no such $n$ exists, we set $\sr(R,I)=\infty$.

The following proposition summarizes the basic properties of stable ranks.
\begin{prop} \label{prop:sr_properties}
Let $R$ be arbitrary commutative unital ring and let $I\trianglelefteq R$ be its ideal.
\begin{itemize}
\item For any ideal $J\trianglelefteq R$, $J\subseteq I$ one has 
\[ \sr(R,I)\leqslant\sr(R),\quad\sr(R/J,I/J)\leqslant\sr(R,I). \]
\item Stable rank of a direct product of rings is equal to the maximum of stable ranks of its factors:
\[ \sr\left(\prod\limits_{i=1}^n R_i\right)=\max\limits_{i=1}^n\left(\sr(R_i)\right). \]
\item Stable rank does not change after taking quotient modulo the nil-radical: $\sr(R)=\sr(R/\Rad(R))$.
\end{itemize}
\end{prop}
\begin{proof} See~\cite{Va71}. \end{proof}

\begin{example}
Since the stable rank of a field is one, one can conclude from the previous proposition that $\sr(R)=1$ for any semilocal ring $R$.
\end{example}
\begin{example}
Other examples of rings of stable rank $1$ include boolean rings, the ring of algebraic integers, the ring of entire functions, the disk-algebra.
\end{example}
\begin{example}
If $J$ is the Jacobson radical of $R$, one has $\sr(R,J)=1$.
\end{example}

\subsection{Absolute stable rank}
For a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ denote by $\mathfrak{I}(u)$ the intersection of ideals of $R$ containing $a_1,\ldots,a_n$.
It is easy to see that a row $a\in R^n$ is unimodular if and only if $\mathfrak{I}(u)=R$. 
Clearly, for any $g\in\GL(n,R)$ one has $\mathfrak{I}(u\cdot g)=\mathfrak{I}(u)$.

A ring $R$ is said to satisfy condition $\ASR_{n}(R,I)$ if for any row $a$ such that $a_1-1,\ldots,a_{n+1}\in I$ one can find elements $b_1,\ldots b_n\in R$ such that
\[ \mathfrak{I}(a_1+b_1a_{n+1},\ldots,a_n+b_na_{n+1})^t = \mathfrak{I}(a_1,\ldots,a_{n+1})^t. \]
It is easy to see that $\ASR_m(R,I)$ implies $\ASR_n(R,I)$ for any $n\geqslant m$. 
By definition, the {\it absolute relative stable rank} $\asr(R,I)$ is the smallest natural $n$ such that $\ASR_n(R, I)$ holds. Similarly, we set $\asr(R,I)=\infty$ if no such $n$ exists. When $I=R$ we call $\asr(R, R)$ simply the {\it absolute stable rank} of $R$ and denote it by $\asr(R)$.

Let $R$ be a commutative ring. We denote by $\Max(R)$ its {\it maximum spectrum} i.e. the set of maximal ideals of $R$ equipped with Zariski topology.
The following result gives an upper estimate for the absolute stable rank.~\cite[Theorem~2.3]{EO}, \cite[Theorem~3.7]{MKV}).
\begin{thm}
Let $R$ be a commutative ring whose maximum spectrum $\Max(R)$ can be covered with a finite family of subsets $X_1\ldots, X_m$ such that the topological dimension of every subset $X_i$ does not exceed some number $d$. Then $\asr(A)\leqslant d+1$.
\end{thm}
Clearly one has
\[ \sr(R,I)\leqslant\asr(R,I)\leqslant\asr(R)\leqslant \dim(\Max(R))+1\leqslant\dim(\Spec(R))+1. \]

>>>>

\subsection{Stable rank}
Recall the definition of \emph{stable rank} introduced by H.~Bass and L.~Vaserstein (see~\cite{Ba64, Va69}).
\begin{dfn}
Recall that a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ is called \emph{$I$-unimodular} if elements $a_1-1,a_2,\ldots,a_n$ are contained in $I$ while $a_1,a_2,\ldots,a_n$ generate $R$ as an ideal.
\end{dfn}
A column $b \in R^n$ will be called $I$-unimodular if its transpose $b^t$ is an $I$-unimodular row. We denote the set of all $I$-unimodular rows (resp. columns) by $\Umd(n,R,I)$ (resp. $\Ums(n,R,I)$).
When $R=I$ we refer to $R$-unimodular rows and columns as unimodular.
\begin{lemma}\label{lemma:relstrlemma}
For any $I$-unimodular row $a=(a_1,\ldots,a_n)$ there exists $I$-unimodular column $b=(b_1,\ldots,b_n)^t$ such that $ab=\sum\limits_{i=1}^n a_i b_i=1$.
\end{lemma}
\begin {dfn} An $I$-unimodular row $a=(a_1,\ldots a_{n+1})$ is called \emph{stable} if one can choose $b_1,\ldots,b_n\in R$ such that the row $(a_1+a_{n+1}b_1,\ldots,a_n+a_{n+1}b_n)$ is also $I$-unimodular. If, moreover, elements $b_1,\ldots,b_n$ can be chosen from  $I$ then such $a$ will be called \emph{$I$-stable}. \end{dfn} 
We say that a pair $(R,I)$ satisfies stable range condition $\SR_n(R,I)$ if any $I$-unimodular row of length $n+1$ is stable.
\begin{lemma}\label{lemma:relstrlemma2}\strut
\begin{enumerate}
\item The condition $\SR_n(R,I)$ implies $\SR_m(R,I)$ for any $m\geqslant n$.
\item The condition $\SR_n(R,I)$ implies that any $I$-unimodular row of length $n+1$ is $I$-stable.
\end{enumerate}
\end{lemma}
By definition, the {\it relative stable rank} $\sr(R,I)$ of a pair $(R,I)$ is the smallest natural number $n$ such that condition $\SR_n(R,I)$ holds.
If no such $n$ exists, we set $\sr(R,I)=\infty$.

The following proposition summarizes the basic properties of stable ranks.
\begin{prop} \label{prop:sr_properties}
Let $R$ be arbitrary commutative unital ring and let $I\trianglelefteq R$ be its ideal.
\begin{itemize}
\item For any ideal $J\trianglelefteq R$, $J\subseteq I$ one has 
\[ \sr(R,I)\leqslant\sr(R),\quad\sr(R/J,I/J)\leqslant\sr(R,I). \]
\item Stable rank of a direct product of rings is equal to the maximum of stable ranks of its factors:
\[ \sr\left(\prod\limits_{i=1}^n R_i\right)=\max\limits_{i=1}^n\left(\sr(R_i)\right). \]
\item Stable rank does not change after taking quotient modulo the nil-radical: $\sr(R)=\sr(R/\Rad(R))$.
\end{itemize}
\end{prop}
\begin{proof} See~\cite{Va71}. \end{proof}
\begin{example}
Since the stable rank of a field is one, one can conclude from the previous proposition that $\sr(R)=1$ for any semilocal ring $R$.
\end{example}
\begin{example}
Other examples of rings of stable rank $1$ include boolean rings, the ring of algebraic integers, the ring of entire functions, the disk-algebra.
\end{example}
\begin{example}
If $J$ is the Jacobson radical of $R$, one has $\sr(R,J)=1$.
\end{example}

\subsection{Absolute stable rank}
For a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ denote by $\mathfrak{J}(a)$ the intersection of all maximal ideals of $R$ containing $a_1,\ldots,a_n$.
It is easy to see that a row $a\in R^n$ is unimodular if and only if $\mathfrak{J}(a)=R$. 
Clearly, for any $g\in\GL(n,R)$ one has $\mathfrak{J}(a\cdot g)=\mathfrak{J}(a)$.
\begin{dfn}
We say that a pair $(R,I)$ satisfies $\ASR_n(R,I)$ if it satisfies $\SR(R,I)$ and, moreover, for any row $a=(a_1,\ldots,a_{n+1})\in I^n$ there exist $b_1,\ldots,b_n\in I$ such that
\[\mathfrak{J}(a_1,\ldots,a_{n+1})=\mathfrak{J}(a_1+b_1a_{n+1},\ldots,a_n+b_na_{n+1}).\]
It is easy to see that $\ASR_m(R,I)$ implies $\ASR_n(R,I)$ for any $n\geqslant m$. The \emph{absolute relative stable rank} $\asr(R,I)$ is the smallest natural $n$ such that $\ASR_n(R, I)$ holds. Similarly, we set $\asr(R,I)=\infty$ if no such $n$ exists. When $I=R$ we call $\asr(R, R)$ simply the \emph{absolute stable rank} of $R$ and denote it by $\asr(R)$.
\end{dfn}
The following reformulation of the absolute stable rank condition was noted by A.~Bak.
\begin{lemma}
$R$ satisfies $\ASR_n(R)$ if and only if for any \emph{unimodular} row $(a_1,\ldots,a_n,b,d)\in\Ums(n+2,R)$ there exist $c_1,\ldots,c_n\in R$ such that $(a_1+c_1d,\ldots,a_n+c_nd,b+td)$ is unimodular for any $t\in R$.
\end{lemma}
\begin{proof}
Assume first that $\ASR_n(R)$ holds. Then for the row $(a_1,\ldots,a_n,d)$ there exist $c_1,\ldots,c_n$ such that $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a_1+c_1d,\ldots,a_n+c_nd)$. Therefore $a_1+c_1d,\ldots,a_n+c_nd$ together with $b$ generate the unit ideal. Of course, one could replace $b$ with $b+td$ from the very start.

Now, given an arbitrary row $(a_1,\ldots,a_n,d)\in{}^{n+1}\!R$, consider a unimodular row $(a_1,\ldots,a_n,1,d)\in\Ums(n+2,R)$. There exist $c_1,\ldots,c_n$ such that
\[(a'_1,\ldots,a'_n,1+td)=(a_1+c_1d,\ldots,a_n+c_nd,1+td)\]
is unimodular for any $t\in R$. Suppose there exist a maximal ideal $\mathfrak{m}\trianglelefteq R$ such that $a'_1,\ldots,a'_n\in\mathfrak{m}$ but some of $d,a_i\notin\mathfrak{m}$. Clearly, $d\notin\mathfrak{m}$ (otherwise $a_i=a'_i-c_id\in\mathfrak{m}$). On the other hand, $1+td\notin\mathfrak{m}$, therefore in the residue field $R/\mathfrak{m}$ one has $1+\bar{t}\bar{d}\neq0$ for any $t$, but $\bar{d}\neq0$ and one can take $\bar{t}=-1/\bar{d}$.
\end{proof}
Let $R$ be a commutative ring. We denote by $\Max(R)$ its \emph{maximum spectrum} i.e. the set of maximal ideals of $R$ equipped with Zariski topology.
The following result gives an upper estimate for the absolute stable rank (see~\cite[Theorem~2.3]{EO}, \cite[Theorem~3.7]{MKV}).
\begin{thm}
Let $R$ be a commutative ring whose maximum spectrum $\Max(R)$ can be covered with a finite family of subsets $X_1\ldots, X_m$ such that the topological dimension of every subset $X_i$ does not exceed some number $d$. Then $\asr(A)\leqslant d+1$.
\end{thm}
Clearly one has
\[ \sr(R,I)\leqslant\asr(R,I)\leqslant\asr(R)\leqslant \dim(\Max(R))+1\leqslant\dim(\Spec(R))+1. \]
